---
title: "question_mark"
author: "Annie Johansson"
date: "2024-08-15"
output: html_document
---

```{r Set-up, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(lmtest)
library(ggplot2)
library(data.table)
library(DescTools)
library(ggExtra)
library(cowplot)

# plotting tools
colors = c("grey40", "tomato3", "turquoise4", "tan2")
export_pdf <- function(plot_object, height = 10, width = 12, name = deparse(substitute(plot_object))) {
  # !! need to end the file name with a / !!
  pdf(file = paste0("~/research-collaboration/question-mark/results_11022025/plots/", name, ".pdf"),
      height = height,
      width = width,
      bg = "white")
  print(plot_object)
  dev.off()
}
```


# Load data

```{r}
# check differences in qm between raw and cleaned dataset

load("~/research-collaboration/question-mark/new data/logs_all.Rdata")
source("~/research-collaboration/question-mark/irtree_prep.R")

dt <- as.data.table(logs_all)
rm(logs_all)
dt[answer == "多", question_mark := 1]
dt[answer != "多", question_mark := 0]
dt[, response := ifelse(question_mark == 1, "qm", ifelse(correct_answered == 0, "error", "correct"))]
table(dt$response)

dat_irt[answer == "多", question_mark := 1]
dat_irt[answer != "多", question_mark := 0]
dat_irt[, response := ifelse(question_mark == 1, "qm", ifelse(correct_answered == 0, "error", "correct"))]
table(dat_irt$response)

dt[, in_sub := user_id %in% dat_irt[, unique(user_id)]]
ggplot(dt[, .(mean_qm = mean(question_mark)), .(in_sub, user_id)]) +
  geom_histogram(aes(x = mean_qm,), bins = 50) +
  facet_wrap(~in_sub, scales = "free") +
  theme_bw()

dt[, mean(question_mark), in_sub]
dat_irt[, mean(q)]

# clean up grade
dt <- dt[grade > 2 & grade < 9]
```

# Descriptives
## Raw data: N qm per diff level
```{r}
dt[, diff_factor := factor(difficulty, levels = (0:2), labels = c("Easy", "Medium", "Hard"))]
dt_diff <- dt[, .(mean_qm = mean(question_mark),
                  sd_qm = sd(question_mark), 
                  ci_l = t.test(question_mark)$conf.int[1], 
                  ci_u = t.test(question_mark)$conf.int[2]), diff_factor]
dt_diff

ggplot(dt_diff, aes(x = diff_factor, y = mean_qm)) +
  geom_bar(stat = "identity", fill = "gray") +
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u), color = "gray30", width = 0.1) +
  labs(x = "Difficulty Level", y = "Proportion QM Response") +
  theme_bw() +
    theme(
    text = element_text(size = 14)
  )
  
```

## Raw data: N qm per grade

```{r}
dt_grade <- dt[, .(mean_qm = mean(question_mark),
                   sd_qm = sd(question_mark),
                  ci_l = t.test(question_mark)$conf.int[1], 
                  ci_u = t.test(question_mark)$conf.int[2]), .(grade)][order(grade)]
dt_grade$grade <- factor(dt_grade$grade, levels = c(3:8))
dt_grade
ggplot(dt_grade, aes(x = grade, y = mean_qm)) +
  geom_path(group = 1, color = "gray") +
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u), color = "gray30", width = 0.1) +
  labs(x = "Grade", y = "Proportion QM Response") +
  theme_bw() +
  theme(
    text = element_text(size = 14)
  )
  
```

## Raw data: N qm per diff level and graade

```{r}
dt_grade_diff <- dt[, .(mean_qm = mean(question_mark),
                  ci_l = t.test(question_mark)$conf.int[1], 
                  ci_u = t.test(question_mark)$conf.int[2]), .(grade, difficulty = diff_factor)][order(grade, difficulty)]
dt_grade_diff$grade <- factor(dt_grade_diff$grade, levels = c(3:8))

ggplot(dt_grade_diff, aes(x = grade, y = mean_qm, fill = difficulty, group = difficulty, color = difficulty)) +
  geom_ribbon(aes(ymin = ci_l, ymax = ci_u), color = NA, alpha = 0.3, width = 0.1) +
  geom_path(linewidth = 1) +
  scale_discrete_manual(aesthetics = c("fill", "color"), values = colors[c(3, 4, 2)], name = "Difficulty Level") +
  labs(x = "Grade", y = "Proportion QM Response") +
  theme_bw() +
    theme(
    text = element_text(size = 14)
  )
```

## Response freq. over rt

```{r}
dt[, t := round(response_in_milliseconds/1000, 1)]

library(DescTools)
dt_t <- dt[t < 20, Freq(response), t][order(t)]
time_plot <- ggplot(dt_t, aes(x = t, y = perc, group = level, color = level)) +
  geom_line(linewidth = 1) +
  geom_point(aes(size = freq), alpha = 0.6) +
  scale_discrete_manual(aesthetics = c("fill", "color"), values = colors[c(1, 2, 4)], name = "Response", labels = c("Correct", "Error", "Question Mark")) +
  geom_vline(xintercept = 20, linetype = "dotted" ) +
  labs(x = "Response Time", y = "Probability of Response", size = "Frequency") +
  theme_bw() +
    theme(
    text = element_text(size = 14)
  )
time_plot

# Add a histogram of rt to the background of the line plot
time_plot_dist <- ggplot() +
  geom_histogram(data = dt[t <= 21], aes(x = t, y = ..density..), bins = 50, color = "gray", fill = "gray", alpha = 0.4) +
  geom_line(data = dt_t, aes(x = t, y = perc, group = level, color = level), linewidth = 1, alpha = 0.6) +
  geom_point(data = dt_t, aes(x = t, y = perc, group = level, color = level, size = freq), alpha = 0.4) +
  scale_discrete_manual(aesthetics = c("fill", "color"), values = colors[c(1, 2, 4)], name = "Response", labels = c("Correct", "Error", "Question Mark")) +
  geom_vline(xintercept = 20, linetype = "dotted") +
  labs(x = "Response Time", y = "Probability of Response", size = "Frequency") +
  theme_bw() +
  theme(
    text = element_text(size = 14)
  )

time_plot_dist
```

## Resp freq. over session count

```{r}
dt_session <- dt[new_user_domain_modified_count > 0 & new_user_domain_modified_count <= 250, .(mean_qm = mean(question_mark),
                  ci_l = t.test(question_mark)$conf.int[1], 
                  ci_u = t.test(question_mark)$conf.int[2], .N), new_user_domain_modified_count][order(new_user_domain_modified_count)]

ggplot(dt_session, aes(x = new_user_domain_modified_count, y = mean_qm)) +
  geom_ribbon(aes(ymin = ci_l, ymax = ci_u), alpha = 0.6, fill = colors[4]) +
  geom_path(group = 1, color = colors[4]) +
  geom_point(aes(size = N), color = colors[4], alpha = 0.6) +
  theme_bw() +
    theme(
    text = element_text(size = 14)
  )

dt_session <- dt[new_user_domain_modified_count > 0 & new_user_domain_modified_count < 100, Freq(response), new_user_domain_modified_count][order(new_user_domain_modified_count)]
ggplot(dt_session[level != "correct"], aes(x = new_user_domain_modified_count, y = perc, group = level, color = level)) +
  geom_line(linewidth = 1) +
  geom_point(aes(size = freq), alpha = 0.6) +
  scale_discrete_manual(aesthetics = c("fill", "color"), values = colors[c(2, 4)], name = "Response", labels = c("Error", "Question Mark")) +
  labs(x = "Session Count", y = "Probability of Response", size = "Frequency") +
  scale_x_continuous(breaks = c(1, 25, 50, 75, 100), limits = c(1, 100)) +
  theme_bw() +
    theme(
    text = element_text(size = 14)
  )

```

# Model analysis

## Load tree model
```{r}
load("~/research-collaboration/question-mark/results_11022025/mod_tree.Rdata")
load("~/research-collaboration/question-mark/results_11022025/sum_mod_tree.Rdata")
```

## Fixed effects

```{r}
int <- sum$coefficients[1]
1 / (1 + exp(-int))

```


## Inspect random effects

```{r, eval = FALSE}
ranefs_mod <- ranef(mod_tree)
# save(ranefs_mod, file = "ranefs_mod.Rdata")
# load("ranefs_mod.Rdata")
hist(plogis(ranefs_mod$user_id[,1]), breaks = 50, main = "Distribution of random user effects, node 1")
hist(plogis(ranefs_mod$user_id[,2]), breaks = 50,main = "Distribution of random user effects, node 2")
plot(plogis(ranefs_mod$user_id[,1]), plogis(ranefs_mod$user_id[,2]), xlab = "Random user effects, node 1", ylab = "Random user effects, node 2")

hist(plogis(ranefs_mod$item_id[,1]), breaks = 50, main = "Distribution of random item effects, node 1")
hist(plogis(ranefs_mod$item_id[,2]), breaks = 50,main = "Distribution of random item effects, node 2")
plot(plogis(ranefs_mod$item_id[,1]), plogis(ranefs_mod$item_id[,2]), xlab = "Random item effects, node 1", ylab = "Random item effects, node 2")
```

## Extract random effects

```{r}
ranef_user_node1 <- ranefs_mod$user_id[,1]
ranef_user_node2 <- ranefs_mod$user_id[,2]

ranef_item_node1 <- ranefs_mod$item_id[,1]
ranef_item_node2 <- ranefs_mod$item_id[,2]

ranefs <- data.frame(node = factor(c(rep(c(1, 2), each = length(ranef_user_node1)), rep(c(1, 2), each = length(ranef_item_node1)))),
                     estimate = c(rep("user", times = length(ranef_user_node1)*2), rep("item", times = length(ranef_item_node1)*2)),
                     id = c(rep(rownames(ranefs_mod$user_id), 2), rep(rownames(ranefs_mod$item_id), 2)),
                     # values are converted to probabilities
                     value = c(ranef_user_node1, ranef_user_node2, ranef_item_node1, ranef_item_node2))
rm(ranef_user_node1, ranef_user_node2, ranef_item_node1, ranef_item_node2)
# save(ranefs, file = "ranefs_df.Rdata")
# load("ranefs_df.Rdata")
```

## Plot distribution of random effects 

```{r}
ggplot(ranefs, aes(x = value, fill = node, color = node)) +
  geom_density(alpha = 0.7) +
  scale_discrete_manual(aesthetics = c("fill", "color"), values = c("tan1", "steelblue")) +
  facet_grid(node~estimate) +
  theme_bw()
```

## Plot correlation between random effects

```{r}
cor_plot_user <- ranefs %>%
  filter(estimate == "user") %>%
  pivot_wider(names_from = node, values_from = value) %>% 
  ggplot(aes(x = `1`, y = `2`)) + 
  geom_point(color = "gray", alpha = 0.3) +
  geom_smooth(method = "lm", color = "tan2", fill = "tan2") +
  labs(x = "Node 1", y = "Node 2") +
  theme_bw() +
  theme(legend.position = "none", text = element_text(size = 12))

cor_plot_item <- ranefs %>%
  filter(estimate == "item") %>% 
  pivot_wider(names_from = node, values_from = value) %>% 
  ggplot(aes(x = `1`, y = `2`)) +
  geom_point(color = "gray", alpha = 0.3) +
  geom_smooth(method = "lm", color = "tomato3", fill = "tomato3") +
  labs(x = "Node 1", y = "Node 2") +
  theme_bw() +
  theme(legend.position = "none", text = element_text(size = 12))

# add marginal dist
cor_plot_item_marginal <- ggMarginal(cor_plot_item + labs(x = "Item Skipping Threshold", y = "Item Difficulty"), 
                                     type = "density", fill = "tomato3", alpha = 0.4, color = "tomato3")
cor_plot_user_marginal <- ggMarginal(cor_plot_user + labs(x = "User Propensity to Skip an Item", y = "User Propensity for Incorrect Response"), 
                                     type = "density", fill = "tan2", alpha = 0.4, color = "tan2")

combined_plot <- plot_grid(cor_plot_item_marginal, 
                           cor_plot_user_marginal, 
                           ncol = 2)
rm(cor_plot_item, cor_plot_item_marginal, cor_plot_user_marginal)
combined_plot
export_pdf(combined_plot, height = 4, width = 6)
```


## Individual response patterns

```{r }
#load("ranefs_mod.Rdata")
#load("logs_clean.Rdata")

df_user <- data.frame()
ind_plots <- list()
labels <- LETTERS[1:8]  
label_index <- 1       

for (i in 1:4) {
  # set boundary for which areas of correlation plot to sample from 
  # upper right, lower right, lower left, upper left
  min_node1 <- c(1, 1, -4, -4)[i]
  max_node1 <- c(6, 6, -1, 0)[i]
  min_node2 <- c(1.5, -2, -2, 1)[i]
  max_node2 <- c(4, 0, 0, 4)[i]
  
  # sample 2 users per quadrant
  set.seed(121103961)
  users <- sample(as.numeric(rownames(ranefs_mod$user_id[which(
    ranefs_mod$user_id$`factor(node)1` > min_node1 &
    ranefs_mod$user_id$`factor(node)1` < max_node1 & 
    ranefs_mod$user_id$`factor(node)2` > min_node2 & 
    ranefs_mod$user_id$`factor(node)2` < max_node2),])), 2)
  
  # assign labels to sampled users
  user_labels <- labels[label_index:(label_index + 1)]
  label_index <- label_index + 2
  
  # add user data and labels for the correlation plot
  df_user <- rbind(
    df_user,
    data.frame(
      id = users,
      label = user_labels
    )
  )
  
  # prepare data for individual user response time plots
  dat_sub <- logs_clean[logs_clean$user_id %in% users, ]
  dat_sub$response_new <- ifelse(dat_sub$q == 1, "Question Mark", ifelse(dat_sub$correct_answered == 0, "Error", "Correct"))
  dat_sub <- dat_sub[order(dat_sub$id), ]
  dat_sub$item_number <- 1:nrow(dat_sub)
  
  # facet labels
  custom_labels <- setNames(paste(user_labels, c(": Low Accuracy, High Skipping",
                                                 ": High Accuracy, High Skipping",
                                                 ": High Accuracy, Low Skipping",
                                                 ": Low Accuracy, Low Skipping")[i]), users)
  
  # individual user response time plots
  p <- ggplot(dat_sub, aes(x = factor(item_number), y = rt, color = response_new)) + 
    geom_path(group = 1, color = "black", alpha = 0.3) + 
    geom_point() + 
    labs(x = "Item count", y = "Response Time", color = "Response") +
    scale_color_manual(values = c("Correct" = "gray40", "Error" = "tomato3", "Question Mark" = "tan1")) +
    facet_wrap(~user_id, nrow = 10, scales = "free", labeller = labeller(user_id = custom_labels)) + 
    scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = length(x)/10)]) +
    theme_bw()
  
  ind_plots[[paste0("plot_", i)]] <- p
}
```


```{r Final plot: individual response patterns, }
ranefs[ranefs$id %in% df_user$user_id,]
df_ind <- merge(ranefs, df_user, by = "id", all.x = TRUE)
df_ind <- df_ind %>%
  filter(estimate == "user") %>%
  pivot_wider(names_from = node, values_from = value) 

df_ind_sub <- df_ind[complete.cases(df_ind$label),]

cor_plot <- ggplot(df_ind, aes(x = `1`, y = `2`)) + 
  geom_point(color = "gray40", alpha = 0.3) +
  geom_smooth(method = "lm", color = "gray7", fill = "gray7") +
  geom_point(data = df_ind_sub, aes(x = `1`, y = `2`), color = "tomato2") +
  geom_label(aes(label = label), size = 3, nudge_x = 0.2, nudge_y = 0.2) +
  labs(x = "User Propensity to Skip an Item", y = "User Propensity for Incorrect Response") +
  theme_bw() +
  theme(legend.position = "none", text = element_text(size = 12))

plotsAB <- plot_grid(ind_plots$plot_1 + ylim(0, 21) + theme(legend.position = "none", axis.title = element_blank()), 
          ind_plots$plot_2 + ylim(0, 21) + theme(legend.position = "none", axis.title = element_blank()), 
          ncol = 1)

plotsCD <- plot_grid(ind_plots$plot_3+ ylim(0, 21) + theme(legend.position = "none", axis.title = element_blank()), 
          ind_plots$plot_4+ ylim(0, 21) + theme(legend.position = "none", axis.title = element_blank()), 
          ncol = 1)

legend_plot <- ggplot(data.frame(x = c(1, 2, 3), y = c(1, 2, 3), label = c("Correct", "Incorrect", "Question Mark")),
                      aes(x = x, y = y, color = label)) +
  geom_point(alpha = 0, size = 4) +
  scale_color_manual(name = "Response Type", 
                     values = c("Correct" = "gray30", "Incorrect" = "tomato3", "Question Mark" = "tan2"),
                     labels = c("Correct", "Incorrect", "Question Mark")) +
  guides(color = guide_legend(override.aes = list(size = 5, alpha = 1))) +  # Control legend appearance
  theme_void() + # Remove axes and background
  theme(legend.position = "bottom", text = element_text(size = 14))


full_grid = plot_grid(plotsCD, 
                      cor_plot + theme(plot.margin = unit(c(2, 0.5, 2, 0.5), "cm")),
                      plotsAB, 
                      ncol = 4,  # Adjust number of columns
                      rel_widths = c(2, 1, 2, 0.5), # Set the width of the legend column
                      rel_heights = c(2, 1, 2))

full_grid_legend = plot_grid(full_grid, 
                             legend_plot + theme(plot.margin = unit(c(0, 2, 0, -1), "cm")), 
                             ncol = 1, rel_heights = c(12, 1))

full_grid_legend
#export_pdf(full_grid_legend, height = 6, width = 18)
```
